---
title: "Covid91 vaccine study Final2023"
author: "Joel Calderin"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex == 'M')
Females = filter(FinalData, FinalData$sex == 'F')
LGBTQ1 = filter(FinalData, FinalData$LGBTQ == 'gay')
DruggiesY = filter(FinalData, FinalData$DrugUser == 'yes')
 
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

### Is the new COVID921 vaccine effective for the Male subgroup?

$H_0$ There is NOT a significant difference in the effectiveness of the COVID921 vaccine for the Male population.

$H_A$ There is a significant difference in the effectiveness of the COVID921 vaccine for the Male population.


## Methods

Both variables, sex (Males) and treatment, are categorical. Since they are both categorical, we will use the Cat~Cat rmd helper file.The COVID921 effectiveness will be our response variable. 


## Results



## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males, type="percent")
```


## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
```


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Conclusion (Males)



# Females

### Is the new COVID921 vaccine effective for the Female subgroup?

$H_0$ There is NOT a significant difference in the effectiveness of the COVID921 vaccine for the female population.

$H_A$ There is a significant difference in the effectiveness of the COVID921 vaccine for the female population.



## Methods

Both variables, sex (Females) and treatment, are categorical. Since they are both categorical, we will use the Cat~Cat rmd helper file.The COVID921 effectiveness will be our response variable. 


## Results



## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females, type="percent")
```


## Numerical Results


```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
```



## Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Conclusion (Females)




# LGBTQ

### Is the new COVID921 vaccine effective for the LGBTQ subgroup?

$H_0$ There is NOT a significant difference in the effectiveness of the COVID921 vaccine for the LGBTQ population.

$H_A$ There is a significant difference in the effectiveness of the COVID921 vaccine for the LGBTQ population.

## Methods

Both variables, LGBTQ and treatment, are categorical. Since they are both categorical, we will use the Cat~Cat rmd helper file.The COVID921 effectiveness will be our response variable. 


## Results



## Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ1, type="percent")
```



## Numerical Results


```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ1)
rowPerc(table1)
```


## Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Conclusion (LGBTQ)



# Druggies

### Is the new COVID921 vaccine effective for the Druggies subgroup?

$H_0$ There is NOT significant difference in the effectiveness of the COVID921 vaccine for the Druggies subgroup.

$H_A$ There is a significant difference in the effectiveness of the COVID921 vaccine for the Druggies subgroup.

## Methods

Both variables, Drug User and treatment, are categorical. Since they are both categorical, we will use the Cat~Cat rmd helper file.The COVID921 effectiveness will be our response variable. 


## Results



## Graphical Results

```{r}
barchartGC(~infected + treatment,data=DruggiesY, type="percent")
```

People who said YES to being drug users. 



## Numerical Results


```{r}
table1 <- xtabs(~infected + treatment, data=DruggiesN)
rowPerc(table1)
table1 <- xtabs(~infected + treatment, data=DruggiesY)
rowPerc(table1)
```


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Conclusion (Druggies)



# Overall Results and Conclusions